How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-06-07 research/paper/draft.tex
reviving the paper, starting with an outline

%\documentclass[iop]{emulateapj}
\documentclass[preprint]{aastex}
%\documentclass[12pt, onecolumn]{emulateapj}
%\documentstyle[aas2pp4,natbib209]{article}

@@

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\textul}{\underline}

@@

This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model of hierarchical inference.  We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.  CHIPPR yields an accurate characterization of $n(z)$ containing information beyond the best-fit estimator produced by traditional procedures.  The publicly available code is easily extensible to other one-point statistics that depend on redshift.

@@

The redshift distribution $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.  Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey, biasing recovery of the cosmological parameters.  For example, if the $n(z)$ used in analysis is offset from the true $n(z)$ by a small, negative constant, it results in underestimating $w_{0}$ and overestimating $\sigma_{8}$ \citep{samuroff_simultaneous_2017}.

Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys including DES, LSST, Euclid, and WFIRST seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.  For decades, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers, as well as other systematics imparted by the properties of the survey, data reduction pipeline(s), and assumptions underlying the analysis \citep{baum_photoelectric_1962}.  These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these nontrivial uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate \citep{koo_overview_1999}.  Such data products have been commonly released by photometric galaxy surveys since SDSS DR7 \citep{abazajian_seventh_2009} using a great variety of methods.

Methods for using photo-$z$ PDFs in cosmological inference remain underdeveloped, with many survey pipelines reducing them to familiar point estimators that are compatible with existing technology or engaging with them heuristically in a manner inconsistent with their probabilistic nature.  Stacking photo-$z$ PDFs, or other mathematically equivalent methods, to obtain an estimator of $n(z)$ is especially popular \citep{cunha_estimating_2009, sheldon_photometric_2012}.  However, photo-$z$ PDFs are probabilistic data products that must be handled in a fully consistent, probabilistic manner; hierarchical inference is the only mathematically valid way to do inference with photo-$z$ PDFs and other probabilistic quantities.

It is desirable to create rigorous methods for using photo-$z$ PDFs in inference, beginning with the simplest one-point statistic of redshift, $n(z)$.  In the spirit of \citet{hogg_inferring_2010, foreman-mackey_exoplanet_2014}, this paper derives a mathematically rigorous approach to inferring $n(z)$ in Sec. \ref{sec:method}.  The presentation of a public code implementing this novel technique is given in Sec. \ref{sec:model_specifics}.  The code is validated on mock data and tested on a subset of BOSS DR10 data in Sec. \ref{sec:experiments}.  We conclude and discuss future directions in Sec. \ref{sec:discussion}.

@@

The redshift distribution $n(z)$ may be understood as the probability of finding a galaxy at a redshift $z$.  We may express $n(z)$ in terms of some functional form $f_{\vec{\theta}}(z)$ with parameters comprising $\vec{\theta}$ according to Eq. \ref{eq:parametrization}.  
\begin{align}
\label{eq:parametrization}
n(z) &\equiv f_{\vec{\theta}}(z)\ \equiv\ p(z|\vec{\theta})
\end{align}
In this work, we wish to characterize the posterior distribution $p(\vec{\theta}|D)$ of the parameters contained in $\vec{\theta}$ defining $n(z)$ given all available data $D$.  The likelihood $p(z|\vec{\theta})$ represents the probability of a random variable, in this case redshift $z$, given the parameters in $\vec{\theta}$ defining the distribution $n(z)$ from which it is drawn.  Our model for inferring $p(\vec{\theta}|D)$ will be introduced in Sec. \ref{sec:derivation}, and alternative methods of estimating $\vec{\theta}$ will be presented in Sec. \ref{sec:alternatives}.

\subsection{Model Specifics}
\label{sec:derivation}

Inference of $p(\vec{\theta}|D)$ is perfectly suited to a hierarchical Bayesian model.  In this problem, the available data $D$ is a catalog of photometry $\{\vec{d}_{i}\}$ of individual galaxies $i$ where each $\vec{d}_{i}$ is a random variable that is a function of its redshift parameter $z_{i}$; the redshift parameters $\{z_{i}\}$ are independently drawn from a function defined by hyperparameters in $\vec{\theta}$.  These physical relationships may be illustrated in Fig. \ref{fig:pgm} by a directed acyclic graph representing a probabilistic graphical model (PGM).
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{pgm.png}
\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown in circles, with observed variables shown in shaded circles.  Relationships between variables are indicated by arrows.  The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.  The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.  Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
\label{fig:pgm}
\end{center}
\end{figure}
This PGM may be used as the basis for a derivation of the desired hyperposterior $p(\vec{\theta}|\{\vec{d}_{i}\})$, assuming our physical model is complete.  From this point on, we will work solely with log probabilities.

We begin with Eq. \ref{eq:bayes} by applying Bayes' Rule to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$.
\begin{align}
\label{eq:bayes}
\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]
\end{align}
To do this, we will choose a log-hyperprior distribution $\ln\left[p(\vec{\theta})\right]$ representing our beliefs about the distribution of the hyperparameters $\vec{\theta}$ in terms of fixed variables that we will assume are known.

The log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$ contains no explicit reference to redshifts, therefore, we employ the PGM to write the marginalization over the redshifts in Eq. \ref{eq:marginalization}.
\begin{align}
\label{eq:marginalization}
\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right] &= \ln\left[\int\ p(\{\vec{d}_{i}\}|\{z_{i}\})\ p(\{z_{i}\}|\vec{\theta})\ d\{z_{i}\}\right]
\end{align}
We shall handle the two terms in the integral separately.

The first term is the likelihood of all galaxy photometry given all galaxy redshifts.  If we assume independence of galaxy photometry, such that $p(\vec{d}_{i}|\{\vec{d}_{i'\neq i}\}, \{z_{i}\})=p(\vec{d}_{i}|z_{i})$, then we may write Eq. \ref{eq:independence}.
\begin{align}
\label{eq:independence}
\ln\left[p(\{\vec{d}_{i}\}|\{z_{i}\})\right] &= \sum_{i}\ln\left[p(\vec{d}_{i}|z_{i})\right]
\end{align}
The second term may also be expanded by assuming that all galaxy redshifts are drawn independently from $n(z)$.
\begin{align}
\label{eq:poisson}
\ln\left[p(\{z_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[p(z_{i}|\vec{\theta})\right]
\end{align}

Photo-$z$ PDFs, on the other hand, are posteriors, probabilities of a parameter, in this case redshift $z$, given the random variable derived from the parameter, in this case photometric data $\vec{d}$.  Though they are commonly written as $p(z)$, a photo-$z$ PDF for galaxy $i$ in the catalog is truly a posterior $p(z_{i}|\vec{d}_{i})$ where $\vec{d}_{i}$ represents photometric data.

This approach makes several assumptions that must be explicitly stated.
\begin{enumerate}
\item The PGM is an expression of our beliefs about the physics of the problem, and the inference will only be valid to the degree that the model is complete.
\item We must choose a hyperprior $p(\vec{\theta})$; though we aim to choose a sufficiently general hyperprior that will not be a dominant source of information in the hyperposterior, it must still be assumed that whatever hyperprior is chosen will be accurate.
\item We have assumed independence of galaxies in our catalog such that each galaxy's photometry $\vec{d}_{i}$ is independent from all other galaxies' photometry $\{\vec{d}_{i'\neq i}\}$ and all other galaxies' redshifts $\{z_{i'\neq i}\}$.
\end{enumerate}

\section{Alternative methods}
\label{sec:alternatives}

\section{Experiments}
\label{sec:experiments}

\section{Discussion}
\label{sec:discussion}

%\appendix{}

\begin{acknowledgements}
AIM thanks Phil Marshall for advising on the production of usable code, Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
\end{acknowledgements}

>>>>

\documentclass[preprint]{aastex}

@@

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\textul}{\underline}
\newcommand{\chippr}{\texttt{CHIPPR} }

@@

This paper presents a novel approach to estimating the posterior distribution 
over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic 
graphical model of hierarchical inference.  We present the Cosmological 
Hierarchical Inference with Probabilistic Photometric Redshifts (\chippr) code 
implementing this technique, as well as its validation on mock data and testing 
on the \texttt{Buzzard} simulations.  \chippr yields a more accurate 
characterization of $n(z)$ containing information beyond the best-fit estimator 
produced by traditional procedures.  The publicly available code is easily 
extensible to other one-point statistics that depend on redshift.

@@

\chippr aims to answer the following questions:

\begin{itemize}
	\item How are photo-$z$ PDFs currently used in cosmology?
	\item Why should we challenge existing methods?
	\item How should photo-$z$ PDFs be used in inference?
	\item How does the result of \chippr compare to established estimators 
of $n(z)$?
\end{itemize}

@@

\subsection{Model}
\label{sec:model}

This is where I will present the DAG and corresponding math, along with an 
explicit discussion of the assumptions that must be satisfied in order for the 
\chippr approach to be valid.

\subsection{Implementation}
\label{sec:implementation}

I will describe the \chippr code here.

\subsection{Metrics}
\label{sec:metrics}

I will discuss the metrics that I use in \ref{sec:validation} and include in 
the release version.

\section{Validation}
\label{sec:validation}

This is where the setup and results of the simulated test cases will be 
presented.

\subsection{Variations on Photo-$z$ Likelihoods}
\label{sec:likelihoods}

I think this is too many test cases!

\begin{itemize}
	\item A fiducial case of Gaussian likelihoods with narrow, 
redshift-independent intrinsic scatter and no other systematics; also the 
fiducial case with larger intrinsic scatter
	\item The fiducial case with catastrophic outliers as seen in 
template-based photo-$z$ estimators (attractor in the space of $z_{phot}$: some 
galaxies at a range of $z_{spec}$ map to the same $z_{phot}$ because their 
shared SED type does not have sufficiently strong features, leading galaxies of 
that type at many $z_{spec}$ to have the same colors)
	\item The fiducial case with catastrophic outliers as seen in 
training-based photo-$z$ estimators (attractor in the space of $z_{spec}$: some 
galaxies at a particular $z_{spec}$ map to a range of $z_{phot}$ because their 
shared SED features fall between filters, leading to many galaxies at 
particular $z_{spec}$ to have the same colors)
\end{itemize}

\subsection{Variations on Interim Priors}
\label{sec:priors}

\subsection{Variations on True Hyperparameters}
\label{sec:truth}

\section{Application}
\label{sec:application}

I will apply \chippr to a realistic simulated dataset (probably 
\texttt{Buzzard}) so the results can be compared to the truth.  The photo-$z$ 
PDFs should be derived using the \texttt{SDSS}-approved $k$ nearest neighbors 
algorithm because it has the most straightforward interim prior.

\section{Conclusion}
\label{sec:conclusion}

I will review the questions from the introduction:

\begin{itemize}
	\item Existing methods have shortcomings that propagate to inaccuracies 
in characterizing the cosmological parameters.\\
	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
in a mathematically consistent manner; \chippr is one such method, conditioned 
on some assumptions.
	\item In addition to coming with its own error distribution, the $n(z)$ 
estimator produced by \chippr is more accurate than established estimators; 
furthermore, propagation of the \chippr result leads to a quantifiable 
improvement in the constraints on cosmological parameters.
\end{itemize}

I may briefly discuss some obvious extensions of \chippr, such as a fully 
hierarchical inference of the cosmological parameters that avoids tomographic 
binning based on photo-$z$ point estimates.

\begin{acknowledgements}
AIM thanks Phil Marshall for significant coaching on code development and paper 
preparation, Mohammadjavad Vakili for insightful input on statistics, Geoffrey 
Ryan for assistance in debugging, and Boris Leistedt for helpful comments 
provided in the preparation of this paper.
\end{acknowledgements}
